---
title: "Achilles Characterization"
author: "DSP - Inspire Mental Health"
date: "`r Sys.Date()`"
output:
  word_document: 
    number_sections: yes
  pdf_document:
    number_sections: yes
  html_document:
    toc: yes
    toc_depth: 4
    number_sections: yes
---

# Clean the Environment

```{r clean environment, results="hide", include=FALSE}
### Start with a clean environment by removing objects in workspace
rm(list=ls())

```

# Set Chunk requirements

```{r setup, results="hide", include=FALSE}

## Set Chunk requirements
knitr::opts_chunk$set(#include = TRUE,
                      echo=FALSE, message = FALSE, warning = FALSE,
                      fig.width=13, fig.height=7)

```

# Setting Work Directory

```{r setting working directory, include=FALSE, results = "hide"}

### Setting work directory
working_directory <- base::setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
#working_directory <- base::setwd(".")

```

# Creating Data, output and Achilles folders

```{r creating subfolders, include=FALSE, results = "hide"}

mainDir <- base::getwd()
subDir_data <- "data"
subDir_output <- "Output"


data_Dir <- base::file.path(mainDir, subDir_data)
output_Dir <- base::file.path(mainDir, subDir_output)


### create data folder
base::ifelse(!base::dir.exists(data_Dir), base::dir.create(data_Dir), "Sub Directory exists")

### create output folders
base::ifelse(!base::dir.exists(output_Dir), base::dir.create(output_Dir), "Sub Directory exists")


### create Output folder for Achilles Analysis
Achilles_Analysis_Dir <- base::file.path(output_Dir, "Achilles Analysis")

base::ifelse(!base::dir.exists(Achilles_Analysis_Dir), base::dir.create(Achilles_Analysis_Dir),
             "Achilles Analysis Sub Directory exists"
             )

```

# Install required packages (CRAN and Github)

```{r loading relevant packages, include=FALSE, results = "hide"}

## Install required packages

### Install CRAN packages
required_packages <- c("RPostgres", "DBI", "tidyverse", "devtools", "DatabaseConnector", "remotes"
                       )

installed_packages <- required_packages %in% base::rownames(utils::installed.packages())

if (base::any(installed_packages==FALSE)) {
  utils::install.packages(required_packages[!installed_packages]
                          #, repos = "http://cran.us.r-project.org"
                          )
}

### load CRAN libraries
base::invisible(base::lapply(required_packages, library, character.only=TRUE))

### development packages
required_dev_packages <- c("OHDSI/Achilles")
required_dev_name_packages <- stringr::str_extract(required_dev_packages, '\\b[^/]+$')

installed_dev_packages <- required_dev_name_packages %in% base::rownames(utils::installed.packages())

if (base::any(installed_dev_packages==FALSE)) {
  remotes::install_github(required_dev_packages[!installed_dev_packages])
} 

### load development libraries
base::invisible(base::lapply(required_dev_name_packages, library, character.only=TRUE))

```


# Store password credentials in .Renviron

- Store your password credentials for postgres in `.Renviron` file and retrieve with `Sys.getenv()`. 

- There are two ways you can do this: By running `usethis::edit_r_environ()` or `file.edit("~/.Renviron")`.

- After running either the above codes, `.Renviron` file will open. Type your password and equate it to an object name. i.e ***postgres_password=your password here***

- Save and close the file then restart R.


# Connecting to database with local CDM instances

```{r Connecting to local database}

database_name <- "mh_staging" #This will be different depending on how you named your database

con <- DBI::dbConnect(drv = RPostgres::Postgres(),
                      dbname = database_name, 
                      host = 'localhost', 
                      port = 5432, 
                      user = 'postgres',
                      password = Sys.getenv("postgres_password")
                      ) 

print(con)

```

# Extract CDM instances of interest from database

```{r}

#List all schemas in database 
list_all_schemas_cdm <- dbGetQuery(con, "SELECT schema_name FROM information_schema.schemata") #output is df

#Extract schemas of interest with the stated format you saved. We saved ours as this: study_1_cdm_r, study_2_cdm_r.....etc

list_all_schemas_study_cdm <- list_all_schemas_cdm$schema_name[grepl("^study_", list_all_schemas_cdm$schema_name)]

```

# Run Achilles

Automated Characterization of Health Information at Large-Scale Longitudinal Evidence Systems (ACHILLES). Achilles provides descriptive statistics on an OMOP CDM database. ACHILLES currently supports CDM version 5 and above (Use major release number or minor number only e.g. 5, 5.3, 5.4)

```{r}

# Create connection details
  cd_achilles <- DatabaseConnector::createConnectionDetails(
    dbms = "postgresql",
    server = paste0("localhost","/",database_name),
    user = "postgres",
    password = Sys.getenv("postgres_password"),
    port = 5432,
    pathToDriver = base::file.path(data_Dir, "JDBC Driver postgresql") #Make sure you put JDBC drivers in data folder
    )

```


```{r}

achilles_analysis <- sapply(list_all_schemas_study_cdm, function(x){
  
  nn <- x
  
  results_schema <- paste0("results_", nn) #Creating object to match results schemas in database: results_study_1_cdm_r....etc
  
  vocabulary_schema <- "vocabulary" #one common vocabulary schema in database
  
  output_folder <- base::file.path(Achilles_Analysis_Dir, nn) #create output folder for individual studies
  
  #run achilles
  Achilles::achilles(connectionDetails = cd_achilles,
                     cdmDatabaseSchema = nn,
                     resultsDatabaseSchema = results_schema,  #no capital letters- brings issues with postgres
                     vocabDatabaseSchema = vocabulary_schema,
                     sourceName = "",
                     createTable = TRUE,
                     smallCellCount = 5,
                     cdmVersion = "5.4",
                     createIndices = TRUE,
                     numThreads = 1,
                     outputFolder = output_folder,
                     optimizeAtlasCache = FALSE
                     )
  
}, simplify = FALSE
)

```


